---
title: "Tidy Fitting of Fluorescence Light Curve Data in R"
author: "Matthew R Nitschke"
date: "23/03/2021"
---

# Import Packages

```{r}
# Libraries for plotting and being tidy
library(tidyverse) # https://www.tidyverse.org/packages/
library(broom) # https://cran.r-project.org/web/packages/broom/vignettes/broom.html
library(lubridate) # https://lubridate.tidyverse.org/

# Libraries for fitting curves
library(nls.multstart) # https://github.com/padpadpadpad/nls.multstart
library(nlstools)
```

# 1 Data imports

## 1.1 Import and clean PAM data. Calculate fluorescence parameters.

```{r}
meta <- read.table(file="https://raw.githubusercontent.com/gerard-ricardo/data/master/postset%20treat%20mil", header= TRUE,dec=",", na.strings=c("",".","NA"))

fn <- list.files(c("rlc_mil_csvs/", "rlc_ten_csvs/"), pattern = ".csv", full.names = TRUE)

processed <- data.frame()

for(i in 1:length(fn)){
temp_data <- read_csv(fn[i]) %>% # reads the ith file
  select(PAR, starts_with("F")) %>%
  select(-starts_with("Fo")) %>%
  pivot_longer(F1:last_col(), names_to = "parameter_idx", values_to = "values") %>%
  mutate(curve_id = str_remove_all(parameter_idx, "F|Fm|m'|Y|II|\\(|\\)")) %>%
  mutate(parameter = str_remove_all(parameter_idx, "[0-9]")) %>%
  filter(!is.na(values)) %>%
  arrange(curve_id, parameter_idx, PAR) %>%
  select(-parameter_idx) %>%
  pivot_wider(names_from = parameter, values_from = values) %>%
  mutate(file_name = fn[i])
  
  processed <- rbind(processed, temp_data)
}

raw_data <- processed %>%
  group_by(file_name, curve_id) %>%
  mutate(unique_id = cur_group_id()) %>%
  rename(Fm = "Fm'") %>%
  mutate(disc = str_remove_all(file_name, "rlc_mil_csvs/|rlc_ten_csvs/")) %>%
  mutate(disc = word(disc, sep = " ")) %>%
  left_join(., meta) %>%
  filter(!is.na(spp)) %>% # m085 has no meta
  filter(dli != "na") %>% # not sure what these are
  mutate(dli = as.numeric(dli))
```

```{r}
pb_data <- raw_data %>%
  group_by(unique_id) %>%
  mutate(Fm = ifelse(Fm <= F, F+0.01, Fm), # There should not be any Fm values < F
         PAR = ifelse(PAR == 0, 0.001, PAR), # PAR values = 0 may result in infinite values during fitting
         FqFm = (Fm - F)/Fm, # Quantum yield of PSII
         Fo.p = first(F) / (first(FqFm) + (first(F)/Fm)), # Fo'
         onemC = (Fm - F)/(Fm - Fo.p), # [1 - C]
         Fv.p = Fm - Fo.p, # Fv'
         onemQ = (Fv.p/Fm)/first(FqFm)) %>% # [1 - Q]
  mutate(FqFm = case_when(FqFm == 0 ~ 0.001, TRUE ~ FqFm)) %>%
  ungroup()
```

```{r}
# Sanity check: PAR count across dataset for any missing data
par_count <- pb_data %>% 
  group_by(PAR) %>%
  count(PAR, sort = TRUE) # there are 16 curves with a par value of 337.000
```

## QC RAW data

```{r}
pb_clean <- pb_data %>%
  mutate(QC = case_when(unique_id == 9 & PAR == 611 ~ "remove",
                        unique_id == 10 & PAR == 611 ~ "remove", 
                        unique_id == 31 & PAR == 611 ~ "remove",
                        unique_id == 32 & PAR == 611 ~ "remove", 
                        unique_id == 33 & PAR == 611 ~ "remove", 
                        TRUE ~ "keep")) %>% # curves 9, 10, 31, 32, 33 have an anomalous value at PAR = 611, why?
  filter(QC == "keep") %>%
  group_by(unique_id) %>%
  slice(which.max(FqFm):which.min(FqFm)) %>% # very suspicious of curves that bottom out then start to increase later at higher PAR values
  ungroup()

pb_clean %>%
  ggplot(aes(PAR, FqFm)) +
  geom_point(size = 2, shape = 21, fill = "blue", alpha = 0.5) +
  facet_wrap(~ unique_id, ncol = 20) + 
  theme(legend.position = "right", aspect.ratio = 1) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  guides(fill = guide_legend(override.aes = list(size = 3, shape = c(21)))) +
  labs(x = expression(paste("E (", mu,"mol photons m"^-2*" s"^-1*")")), 
       y = 'Fq/Fm (dimensionless) +- SE')
```

# 2. Fitting of light curve data

## 2.1 FqFm vs E (PAR): Define the equation

```{r}
# Define the Hennige et al 2008 FqFm equation
Hennige <- function(FqFmmax, Ek, x) {
  model <- ((FqFmmax*Ek)*(1-exp(-x/Ek)))/x
  return(model)
}
```

## 2.2 Fit the FqFm LC using purrr::map across groups

```{r}
# Fit the quantum yield against the PAR data 
FqFmfits <- pb_clean %>%
  group_by(., unique_id, disc, spp, dli, spec ) %>% # any metadata that needs to be propagated forward
  nest() %>%
  mutate(fit = purrr::map(data, ~ nls_multstart(FqFm ~ Hennige(FqFmmax, Ek, x = PAR),
                     data = .x,
                     iter = 250,
                     start_lower = c(FqFmmax = 0.05, Ek = 5),
                     start_upper = c(FqFmmax = 0.95, Ek = 500),
                     supp_errors = 'Y',
                     convergence_count = 100,
                     na.action = na.omit,
                     lower = c(FqFmmax = 0.05, Ek = 5))))

```

## 2.3 Tidy the fits and generate 95% confidence intervals of parameters

```{r}
# get summary
FqFminfo <- FqFmfits %>%
  mutate(summary = map(fit, glance)) %>%
  unnest(summary)

# get parameters
FqFmparams <- FqFmfits %>%
  mutate(., p = map(fit, tidy)) %>%
  unnest(p)

# get confidence intervals
FqFmCI <- FqFmfits %>%
  mutate(., cis = map(fit, confint2),
         cis = map(cis, data.frame)) %>%
  unnest(cis) %>%
  rename(., conf.low = X2.5.., conf.high = X97.5..) %>%
  group_by(., unique_id) %>%
  mutate(., term = c('FqFmmax', 'Ek')) %>%
  ungroup() %>%
  select(., -data, -fit)

# merge parameters and CI estimates
FqFmparams <- merge(FqFmparams, FqFmCI, by = intersect(names(FqFmparams), names(FqFmCI)))

# Create long PAR list
new_preds <- pb_clean %>%
  do(., data.frame(PAR = seq(min(.$PAR), max(.$PAR), length.out = 200), stringsAsFactors = FALSE))

# Augment predictions from fits
predictions <- FqFmfits %>%
  mutate(., p = map(fit, augment, newdata = new_preds)) %>%
  unnest(p) %>%
  rename(., FqFm = .fitted) %>%
  group_by(unique_id) %>%
  mutate(prediction_id = cur_group_id()) %>%
  select(-fit, -data)
```

## 2.4 Plot of FqFm and fits

```{r}
ggplot() +
  geom_vline(aes(xintercept = estimate), FqFmparams %>% filter(term == "Ek")) +
  geom_rect(aes(xmin = conf.low, xmax = conf.high, ymin = 0, ymax = Inf), fill = "red", alpha = 0.5, FqFmparams %>% filter(term == "Ek")) +
  geom_line(aes(PAR, FqFm, group = prediction_id), col = "black", alpha = 0.5, predictions) +
  geom_point(aes(PAR, FqFm, fill = dli), size = 2, shape = 21, alpha = 0.8, pb_clean) +
  facet_grid(vars(spp), vars(spec)) +
  #scale_x_continuous(trans = 'log10') +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_viridis_c(option = "magma") +
  theme(legend.position = "right", aspect.ratio = 1) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  guides(fill = guide_legend(override.aes = list(size = 3, shape = c(21)))) +
  ggtitle("Fq/Fm versus E with model fit ((FqFmmax*Ek)*(1-exp(-x/Ek)))/x") +
  labs(x = expression(paste("E (", mu,"mol photons m"^-2*" s"^-1*")")), 
       y = 'Fq/Fm (dimensionless) +- SE')
```

## 2.5 Plot the variance of the paramters

```{r}
FqFmparams %>%
  ggplot(aes(dli, estimate)) + 
  geom_point(aes(fill = spec), shape = 21, size = 2) +
  facet_grid(vars(term), vars(spp), scales = "free") +
  theme(aspect.ratio = 1) +
  scale_fill_manual(values = c("white", "yellow"))
```

# 3. Light dependant dynamic quenching

## 3.1 Visualise 1 - C vs 1 - Q

```{r}
pb_clean %>%
  ggplot(aes(onemC, onemQ, group = unique_id)) +
  geom_path(aes(colour = dli)) +
  #geom_point(aes(fill = PAR), shape = 21, size = 3) +
  facet_grid(vars(spp), vars(spec)) +
  #scale_fill_viridis_c(option = "magma") + # Requires viridis package
  #scale_shape_manual(values = c(21, 22)) +
  geom_abline(slope = 1) +
  theme(aspect.ratio = 1)
```